---
title: "McKinsey Jan"
author: "Veronika Tamaio Flores"
date: "1/20/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(car)
library(lubridate)
library(tidyr)
library(randomForest)
library(randomForestExplainer)

```

##Uploading data and having first glance on what we have:
```{r upload data}
data <- read.csv("train.csv")
glimpse(data)
cat("We have", nrow(data), "observations and", ncol(data), "variables in the dataset")
```

It is obvious as well that some of the variables (e.g. DOB or Var1) have improper classes. This will be resolved in the Data Cleaning section.

##Examining the data:
1. Cheking empty values
2. Cheking missing values
3. Checking data consistency
```{r check empty, echo=FALSE}
#checking if there are any empty values for each variable
cat("Empty spaces", colnames(data[1]), ":", sum(data$ID == "", na.rm = T))
cat("Empty spaces", colnames(data[2]), ":", sum(data$Gender == "", na.rm = T))
cat("Empty spaces", colnames(data[3]), ":", sum(data$DOB == "", na.rm = T))
cat("Empty spaces", colnames(data[4]), ":", sum(data$Lead_Creation_Date == "", na.rm = T))
cat("Empty spaces", colnames(data[5]), ":", sum(data$City_Code == "", na.rm = T))
cat("Empty spaces", colnames(data[6]), ":", sum(data$City_Category == "", na.rm = T))
cat("Empty spaces", colnames(data[7]), ":", sum(data$Employer_Code == "", na.rm = T))
cat("Empty spaces", colnames(data[8]), ":", sum(data$Employer_Category1 == "", na.rm = T))
cat("Empty spaces", colnames(data[9]), ":", sum(data$Employer_Category2 == "", na.rm = T))
cat("Empty spaces", colnames(data[10]), ":", sum(data$Monthly_Income == "", na.rm = T))
cat("Empty spaces", colnames(data[11]), ":", sum(data$Customer_Existing_Primary_Bank_Code == "", na.rm = T))
cat("Empty spaces", colnames(data[12]), ":", sum(data$Primary_Bank_Type == "", na.rm = T))
cat("Empty spaces", colnames(data[13]), ":", sum(data$Contacted == "", na.rm = T))
cat("Empty spaces", colnames(data[14]), ":", sum(data$Source == "", na.rm = T))
cat("Empty spaces", colnames(data[15]), ":", sum(data$Source_Category == "", na.rm = T))
cat("Empty spaces", colnames(data[16]), ":", sum(data$Existing_EMI == "", na.rm = T))
cat("Empty spaces", colnames(data[17]), ":", sum(data$Loan_Amount == "", na.rm = T))
cat("Empty spaces", colnames(data[18]), ":", sum(data$Loan_Period == "", na.rm = T))
cat("Empty spaces", colnames(data[19]), ":", sum(data$Interest_Rate == "", na.rm = T))
cat("Empty spaces", colnames(data[20]), ":", sum(data$EMI == "", na.rm = T))
cat("Empty spaces", colnames(data[21]), ":", sum(data$Var1 == "", na.rm = T))
cat("Empty spaces", colnames(data[22]), ":", sum(data$Approved == "", na.rm = T))

```

We can see that there are quite a few empty values in the dataset. Let us continue the examination and then dealing with problems we've found.

```{r check missing, echo=FALSE}
#checking missing values for each variable
cat("NAs:", colnames(data[1]), sum(is.na(data$ID)))
cat("NAs:", colnames(data[2]), sum(is.na(data$Gender)))
cat("NAs:", colnames(data[3]), sum(is.na(data$DOB)))
cat("NAs:", colnames(data[4]), sum(is.na(data$Lead_Creation_Date)))
cat("NAs:", colnames(data[5]), sum(is.na(data$City_Code)))
cat("NAs:", colnames(data[6]), sum(is.na(data$City_Category)))
cat("NAs:", colnames(data[7]), sum(is.na(data$Employer_Code)))
cat("NAs:", colnames(data[8]), sum(is.na(data$Employer_Category1)))
cat("NAs:", colnames(data[9]), sum(is.na(data$Employer_Category2)))
cat("NAs:", colnames(data[10]), sum(is.na(data$Monthly_Income)))
cat("NAs:", colnames(data[11]), sum(is.na(data$Customer_Existing_Primary_Bank_Code)))
cat("NAs:", colnames(data[12]), sum(is.na(data$Primary_Bank_Type)))
cat("NAs:", colnames(data[13]), sum(is.na(data$Contacted)))
cat("NAs:", colnames(data[14]), sum(is.na(data$Source)))
cat("NAs:", colnames(data[15]), sum(is.na(data$Source_Category)))
cat("NAs:", colnames(data[16]), sum(is.na(data$Existing_EMI)))
cat("NAs:", colnames(data[17]), sum(is.na(data$Loan_Amount)))
cat("NAs:", colnames(data[18]), sum(is.na(data$Loan_Period)))
cat("NAs:", colnames(data[19]), sum(is.na(data$Interest_Rate)))
cat("NAs:", colnames(data[20]), sum(is.na(data$EMI)))
cat("NAs:", colnames(data[21]), sum(is.na(data$Var1)))
cat("NAs:", colnames(data[22]), sum(is.na(data$Approved)))

#checking how many complete observations are in the dataset
cat("There are", sum(complete.cases(data)), "complete observations in the dataset")
```

We have quite a few NA's as well. Almost all of them are observed in columns connected to loans.

It is time to check the outliers
```{r check outliers}
summary(data$Monthly_Income)
summary(data$Loan_Amount)
summary(data$Loan_Period)
summary(data$Interest_Rate)
summary(data$Existing_EMI)
summary(data$EMI)
```
We have outliers in a few variables such as Monthly_Income, Existing_EMI and EMI. 

Now let us check the levels of factor variables.
```{r check factor levels}
nlevels(data$Gender)
nlevels(data$City_Category)
nlevels(data$Employer_Category1)
nlevels(data$Customer_Existing_Primary_Bank_Code)
nlevels(data$Primary_Bank_Type)
nlevels(data$Contacted)
nlevels(data$Source)
nlevels(data$Source_Category)
```
We have 8 factor variables with lveles range from 2 for Gender to 58 for Customer_Existing_Primaty_Bank_Code.

## Cleaning the data
During the data examination we found a few problems:
- empty values;
- missing values;
- improper variable's classes;
- outliers.

Let us start to solve this problems:
- assigning proper classes:
```{r data cleaning - classes}
data$ID <- as.character(data$ID)
data$DOB <- dmy(data$DOB)
data$Lead_Creation_Date <- dmy(data$Lead_Creation_Date)
data$City_Code <- as.character(data$City_Code)
data$Employer_Code <- as.character(data$Employer_Code)
```

- dealing with empty values:
```{r data cleaning - empty values}
#Discard empty values for DOB - we only have 15 of them
data <- data %>%
  filter(data$DOB !=is.na(data$DOB))

#Discard empty valuess for City_Code - we only have 814 of them
data <- data %>%
  filter(data$City_Code !="")

#Check if City_Category still have empty values
cat("Empty spaces", colnames(data[6]), ":", sum(data$City_Category == "", na.rm = T))
#As far as there are no more empty values in City_Category, we leave it as it is

#Fill empty values for factors
levels(data$Employer_Category1) <- c("unknown", "A", "B", "C")
levels(data$Customer_Existing_Primary_Bank_Code) <- c("unknown", "B001", "B002", "B003", "B004", "B005", "B006", "B007", "B008", "B009", "B010", "B011", "B012", "B013", "B014", "B015", "B016", "B017", "B018", "B019", "B020", "B021", "B022", "B023", "B024", "B025", "B026", "B027", "B028", "B029", "B030", "B031", "B032", "B033", "B034", "B035", "B036", "B037", "B038", "B039", "B040", "B041", "B042", "B043", "B044", "B045", "B046", "B047", "B048", "B049", "B050", "B051", "B052", "B053", "B054", "B055", "B056", "B057")
levels(data$Primary_Bank_Type) <- c("unknown", "G", "P")
```

- dealing with missing values:
```{r data cleaning - missing values}
#Discard empty values for Employer_Category2 
data <- data %>%
  filter(data$Employer_Category2 !=is.na(data$Employer_Category2))

#Fill empty values for loan connected variables
data$Existing_EMI[is.na(data$Existing_EMI)] <- -1
data$Loan_Amount[is.na(data$Loan_Amount)] <- -1
data$Loan_Period[is.na(data$Loan_Period)] <- -1
data$Interest_Rate[is.na(data$Interest_Rate)] <- -1
data$EMI[is.na(data$EMI)] <- -1
```

- outliers: 
```{r data cleaning - outliers}
#Create function to drop outliers
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}
#Removing outliers
data$Monthly_Income <- remove_outliers(data$Monthly_Income)
data <- data %>%
  filter(data$DOB<Sys.Date())

#Cleaning new NA's
data <- data %>%
  filter(data$Monthly_Income !=is.na(data$Monthly_Income))

#We will not drop any other outliers because they may have influence on our future predictions 
```

##Features engineering
The data is clean now. It is time to create new features.
```{r feature engineering}
#Age
data2 <- data %>%
  mutate(Age=as.numeric(round((Sys.Date()-data$DOB)/365)))
```

Now the dataset is ready for predictions!
```{r clean data, echo=FALSE}
cat("We have", nrow(data2), "observations and", ncol(data2), "variables in the dataset")
```

Creating the model
```{r model}
model <- glm(formula=Approved ~ Monthly_Income + Var1, family=binomial, data=data2)
```

Upload testing file and process it
```{r test cleaning}
test <- read.csv("test.csv")

test$ID <- as.character(test$ID)
test$DOB <- dmy(test$DOB)
test$Lead_Creation_Date <- dmy(test$Lead_Creation_Date)
test$City_Code <- as.character(test$City_Code)
test$Employer_Code <- as.character(test$Employer_Code)

levels(test$Employer_Category1) <- c("unknown", "A", "B", "C")
levels(test$Customer_Existing_Primary_Bank_Code) <- c("unknown", "B001", "B002", "B003", "B004", "B005", "B006", "B007", "B008", "B009", "B010", "B011", "B012", "B013", "B014", "B015", "B016", "B017", "B018", "B019", "B020", "B021", "B022", "B023", "B024", "B025", "B026", "B027", "B028", "B029", "B030", "B031", "B032", "B033", "B034", "B035", "B036", "B037", "B038", "B039", "B040", "B041", "B042", "B043", "B044", "B045", "B046", "B047", "B048", "B049", "B050", "B051", "B052", "B053", "B054", "B055", "B056", "B057")

levels(test$Primary_Bank_Type) <- c("unknown", "G", "P")

test$DOB[is.na(test$DOB)] <- "1990-01-01"

test1 <- test %>%
  mutate(Age=as.numeric(round((Sys.Date()-test$DOB)/365)))

test1$Existing_EMI[is.na(test1$Existing_EMI)] <- -1
test1$Loan_Amount[is.na(test1$Loan_Amount)] <- -1
test1$Loan_Period[is.na(test1$Loan_Period)] <- -1
test1$Interest_Rate[is.na(test1$Interest_Rate)] <- -1
test1$EMI[is.na(test1$EMI)] <- -1
```

Testing model on the data
```{r testing}
Approved <-predict.glm(model, newdata = test1, type="response")

#Creating the solution
ID <- test1 %>%
  select(ID)
solution <- cbind(ID, Approved)

#Writing of solution
write.csv(solution, file="solution.csv", row.names=F)
```

